---
title: "SI"
bibliography: 
  - references.bib
  - references_other.yaml
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: none # Use pandoc default, otherwise csl can't be used for reference formatting
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: yes
    latex_engine: xelatex
    pandoc_args: [ "--csl", "journal-of-plant-research.csl"]
    includes:
      in_header: journal-of-plant-research-si.sty
header-includes: 
  \usepackage{float}
  \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
editor_options: 
  markdown: 
    wrap: sentence
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = FALSE)
# if running interactively, source _targets.R first
# source("_targets.R")
```

```{r load-data}
tar_load(
  names = c(
    "filmy_dt_chamber",
    "filmy_dt"
  )
)
```

```{r load-functions}
# Define functions specific to this file
plot_chamber_rh <- function (data, time_limits) {
  ggplot(data, aes(x = date_time, y = 0.01*rh)) +
    geom_line() +
    scale_x_datetime(limits = time_limits) +
    scale_y_continuous(label = label_percent(accuracy = 1)) +
    labs(y = "Relative humidity") +
    theme(
      legend.position = "none",
      axis.title.x = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank())
}

plot_chamber_temp <- function (data, time_limits) {
  ggplot(data, aes(x = date_time, y = temp)) +
    geom_line() +
    scale_x_datetime(limits = time_limits) +
    scale_y_continuous(label = label_number(suffix = "\u00b0C", accuracy = 1)) +
    labs(y = "Temperature") +
    theme(
      legend.position = "none",
      axis.title.x = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank())
}

plot_sporo_chamber_time <- function (data, time_limits) {
    ggplot(data, aes(x = species, ymin = pre, ymax = post, y = post, group = species_duration, linetype = dry_time)) +
    geom_linerange(position = position_dodge(0.5)) +
    geom_point(position = position_dodge(0.5)) +
    scale_y_datetime(limits = time_limits) +
    scale_linetype_manual(values = c("15" = "solid", "2" = "solid")) +
    coord_flip() +
    labs(x = "Species") +
    theme(
      axis.title.x = element_blank(),
      legend.position = "none"
    )
}

plot_sporo_chamber <- function (sporo_times_data, filmy_dt_chamber_data, salt_select, k_select) {
  
  # Get mean times of each cluster of DT test samples for the selected drying salt
  sporo_clusters <- cluster_sporo_times(sporo_times_data, salt_select, k_select)
  
  # Calculate time range for the clustered data
  sporo_time_range <- c(min = min(sporo_clusters$pre), max = max(sporo_clusters$post))
  
  # Subset DT chamber data
  sporo_chamber_data <-
    filmy_dt_chamber_data %>% 
    filter(date_time >= sporo_time_range[["min"]], date_time <= sporo_time_range[["max"]], salt == salt_select)
  
  # RH plot
  rh_plot <- plot_chamber_rh(sporo_chamber_data, sporo_time_range)
  
  # Temp plot
  temp_plot <- plot_chamber_temp(sporo_chamber_data, sporo_time_range)
  
  # Dry times plot
  time_plot <- plot_sporo_chamber_time(sporo_clusters, sporo_time_range)
  
  # Combine plots
  rh_plot + temp_plot + time_plot + plot_layout(ncol = 1)
}

#' Cluster times from a sporophyte desiccation tolerance (DT) test
#' 
#' Calculates mean pre- and post-DT test times by species and duration of DT test
#' (2 or 15 days), also adds a `species_duration` column to the output for plotting.
#'
#' @param sporo_time_data Dataframe with DT test data for sporophytes
#' @param salt_select Name of salt to subset data
#' @param k_select Number of groups to cluster. For example, if the DT
#' test was done in 3 groups, `k_select` = 3.
#'
#' @return Dataframe
cluster_sporo_times <- function (sporo_time_data, salt_select, k_select) {
  
  sporo_2d_clusters <- 
    sporo_time_data %>%
    filter(salt == salt_select, dry_time == "2") %>%
    add_clusters("pre", k_select) %>% # choose K by inspecting raw data first
    group_by(species, cluster, salt, dry_time) %>%
    summarize(pre = mean(pre), post = mean(post), n = n(), .groups = "drop") # %>%
    # select(-cluster)
  
  sporo_15d_clusters <- 
    sporo_time_data %>%
    filter(salt == salt_select, dry_time == "15") %>%
    add_clusters("pre", k_select) %>%
    group_by(species, cluster, salt, dry_time) %>%
    summarize(pre = mean(pre), post = mean(post), n = n(), .groups = "drop") # %>%
    # select(-cluster)
  
  bind_rows(sporo_2d_clusters, sporo_15d_clusters) %>%
    mutate(species_duration = paste(species, dry_time))
  
}

```

```{r gametophyte-dt-2013, fig.cap = ""}
# Calculate mean start/end times for gametophyte DT groups
gameto_mean_times <-
  filmy_dt %>%
  filter(generation == "gametophyte") %>%
  select(individual, contains("time_")) %>%
  pivot_longer(names_to = "condition", values_to = "date_time", -individual) %>%
  filter(!is.na(date_time)) %>%
  mutate(condition = str_remove_all(condition, "time_") %>%
           str_replace_all("30min", "post")) %>%
  filter(condition %in% c("pre", "post")) %>%
  add_count(individual) %>%
  filter(n > 1) %>%
  select(-n) %>%
  pivot_wider(names_from = "condition", values_from = "date_time") %>%
  # Cluster starts times into groups (here, I know there are 10 groups)
  add_clusters("pre", 10) %>%
  mutate(cluster = as.factor(cluster)) %>%
  group_by(cluster) %>%
  summarize(pre = mean(pre), post = mean(post), n = n(), .groups = "drop")
  
gameto_time_limits <- c(min = min(gameto_mean_times$pre), max = max(gameto_mean_times$post))

# DT start/end times plot
dt_times_gameto_2013_plot <- ggplot(
  gameto_mean_times, 
  aes(x = cluster, ymin = pre, ymax = post, y = post, group = cluster)) +
  geom_linerange(position = position_dodge(0.5)) +
  geom_point(position = position_dodge(0.5)) +
  scale_y_datetime(limits = gameto_time_limits) +
  coord_flip() +
  labs(x = "Group") +
  theme(
    axis.title.x = element_blank()
  )

# Subset DT chamber data
chamber_gameto_2013 <- filmy_dt_chamber %>% filter(generation == "gametophyte", year == "2013")

# RH plot
chamber_gameto_2013_rh_plot <-
  plot_chamber_rh(chamber_gameto_2013, gameto_time_limits)
  
# Temp plot
chamber_gameto_2013_temp_plot <-
  plot_chamber_temp(chamber_gameto_2013, gameto_time_limits)

chamber_gameto_2013_temp_plot + chamber_gameto_2013_rh_plot + dt_times_gameto_2013_plot + plot_layout(ncol = 1)

ggsave(here::here("results/S1_fig.pdf"))
```

**Figure S1**. Conditions inside desiccation chamber and duration of desiccation treatment for gametophytes. MgNO3 was used to maintain constant humidity at \circa{ }58%. All gametophytes were treated to desiccation for 2 days, except for group 8, which was treated for four days by accident.

```{r sporophyte-dt-2013-prep}
# Extract pre- and post-DT treatment times for sporophytes
sporo_times <-
  filmy_dt %>%
  filter(generation == "sporophyte") %>%
  select(species:dataset, contains("time_")) %>%
  pivot_longer(names_to = "condition", values_to = "date_time", contains("time_")) %>%
  filter(!is.na(date_time)) %>%
  mutate(condition = str_remove_all(condition, "time_") %>%
           # Consider first recovery measurement (at 30 min) to be "post-DT treatment"
           str_replace_all("30min", "post")) %>%
  filter(condition %in% c("pre", "post")) %>%
  add_count(species, salt, dry_time, individual, dataset) %>%
  filter(n > 1) %>%
  select(-n) %>%
  pivot_wider(names_from = "condition", values_from = "date_time")
```

```{r sporophyte-dt-2013-LiCl, fig.cap = ""}
# Get mean times of each cluster of DT test samples for the selected drying salt
sporo_clusters_2013 <- cluster_sporo_times(filter(sporo_times, str_detect(dataset, "2013")), "LiCl", 3)
sporo_clusters_2014 <- cluster_sporo_times(filter(sporo_times, str_detect(dataset, "2014")), "LiCl", 1)

sporo_clusters <- bind_rows(sporo_clusters_2013, sporo_clusters_2014)

# Calculate time range for the clustered data
sporo_time_range_2013 <- c(min = min(sporo_clusters_2013$pre), max = max(sporo_clusters_2013$post))
sporo_time_range_2014 <- c(min = min(sporo_clusters_2014$pre), max = max(sporo_clusters_2014$post))

# Subset DT chamber data
licl_chamber_data_2013 <-
  filter(filmy_dt_chamber, 
    date_time >= sporo_time_range_2013[["min"]], 
    date_time <= sporo_time_range_2013[["max"]], 
    salt == "LiCl")
licl_chamber_data_2014 <- 
  filter(filmy_dt_chamber, 
    date_time >= sporo_time_range_2014[["min"]], 
    date_time <= sporo_time_range_2014[["max"]], 
    salt == "LiCl") %>%
  # There are two sets of times: 2014-06-17 to 2014-07-02, and 2014-07-06 to 2014-07-08
  # Add a group for these so ggplot doesn't connect lines betweeen them
  mutate(
    day = yday(date_time),
    group = ifelse(day < 185, "a", "b"))

licl_chamber_data <- bind_rows(licl_chamber_data_2013, licl_chamber_data_2014)
licl_temp_range <- c(min(licl_chamber_data$temp), max(licl_chamber_data$temp))
licl_rh_range <- c(min(licl_chamber_data$rh), max(licl_chamber_data$rh))*0.01

# RH plot
rh_plot_2013 <- plot_chamber_rh(licl_chamber_data_2013, sporo_time_range_2013) +
  scale_y_continuous(limits = licl_rh_range, label = label_percent(accuracy = 1))

rh_plot_2014 <- ggplot(licl_chamber_data_2014, aes(x = date_time, y = 0.01*rh, group = group)) +
  geom_line() +
  scale_x_datetime(limits = sporo_time_range_2014) +
  blank_x_theme() +
  blank_y_theme() +
  theme(legend.position = "none")

# Temp plot
temp_plot_2013 <- plot_chamber_temp(licl_chamber_data_2013, sporo_time_range_2013) +
  scale_y_continuous(limits = licl_temp_range, label = label_number(suffix = "\u00b0C", accuracy = 1))

temp_plot_2014 <- ggplot(licl_chamber_data_2014, aes(x = date_time, y = temp, group = group)) +
  geom_line() +
  scale_x_datetime(limits = sporo_time_range_2014) +
  blank_x_theme() +
  blank_y_theme() +
  theme(legend.position = "none")

# Dry times plot
time_plot_2013 <- plot_sporo_chamber_time(sporo_clusters, sporo_time_range_2013) +
  labs(y = "2013") +
  theme(axis.title.x = element_text()) 
time_plot_2014 <- plot_sporo_chamber_time(sporo_clusters, sporo_time_range_2014)  +
  labs(y = "2014")  +
  theme(axis.title.x = element_text()) +
  blank_y_theme()

# Combine plots
temp_plot_2013 + temp_plot_2014 + 
  rh_plot_2013 + rh_plot_2014 + 
  time_plot_2013 + time_plot_2014 + 
  plot_layout(ncol = 2)

ggsave(here::here("results/S2_fig.pdf"))
```

**Figure S2**. Conditions inside desiccation chamber and duration of desiccation treatment for sporophytes at most intense desiccation level. LiCl was used to maintain constant humidity at \circa{ }18%.
